---
title: "STAR Counts Analysis"
---

```{r}
library(DESeq2)
```

Let's load in the data and set some metadata.

```{r}
cts <- read.csv("../data/processed/star_counts.txt", sep="\t", comment.char='#', row.names="Geneid")[, -1:-5]
colnames(cts)[1] = "WT_NR_A"
colnames(cts)[2] = "WT_NR_B"
colnames(cts)[3] = "WT_CR_A"
colnames(cts)[4] = "WT_CR_B"

coldata <- data.frame(row.names=colnames(cts), "condition"=c("Control", "Control", "Treat", "Treat"))
```

Then we can run the `DESeq` pipeline.

```{r, message=FALSE}
dds <- DESeqDataSetFromMatrix(countData=cts, colData=coldata, design=~condition)
dds <- DESeq(dds)
```

Let's start with looking at what are the top $50$ most differentially expressed genes by p-value.

```{r}
res <- results(dds)
as.data.frame(res[order(res$pvalue),][0:50, ])
```

Ok, let's see log fold change where genes with a $p < 0.05$ are in red.

```{r}
plotMA(res, 0.05, main="Yeast Non-restricted vs Calorie-restricted")
```

Wow! That's a lot of highly-differentially expressed genes!

Let's see how different they look between the two groups with a PCA plot.

```{r}
plotPCA(rlogTransformation(dds, blind=TRUE))
```

Looks like the two groups definitely have a different expression profile!